---
title: "Cx4073 : Assignment 3"
author: "Lim Hyun Jeong, A Priyanka, Vinupriya Subramaniyam"
date: "U1521499J U1522958L, U1522959H"
output:
  html_document:
    theme: united
    highlight: tango
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

### Finding Market Segments in Online Retail

Import the CSV data file `assign1_RetailData.csv` for analysis, and quickly check the structure of the data.

```{r}
retailData <- read.csv("assign3_RetailData.csv", header = TRUE)
str(retailData)
```

The following table summarizes the variables in the dataset. Note the type carefully, and convert if required.

| Variable | Description | Remarks | 
| -------- | ------------------- | ------------------------------------------ |
| InvoiceNo | The serial number on the Invoice | May be different for same product and same customer. |
| StockCode | Product (item) code in the Store | Uniquely assigned to every distinct Product (item). |
| Description | Product (item) name in the Store | May be different for the same StockCode. Do check. |
| Quantity | Product Quantity in the Transaction | Will of course change with different transactions. |
| InvoiceDate | Date-Time of the Transaction | Can you use this too? Try, but it is strictly optional. |
| UnitPrice | Product Price per Unit (in Dollars) | Should be the same for the same Product. Do check. |
| CustomerID | Customer code in the Store | Same customer may make various purchases. Check carefully. |
| Country | Country of Residence of Customer | Should be the same for each individual Customer. |

Hint : Think of creating a Customer-Product matrix before you start your market segmentation. First step is deciding the elements in the matrix, and the second step if to choose the notion of distance that whould be appropriate for this case. See if you can effectively use the variables like `Quantity`, `UnitPrice`, `InvoiceDate` and `Country` in this market segmentation. Don't forget to interpret the market segments that you get!

---

**Continue with Data Exploration, creation of Customer-Product matrix, choice of Distance, various Clustering algorithms, and Interpretation. Submit this .Rmd file as your Solution : [StudentID1_StudentID2_StudentID3].Rmd, where the three StudentIDs are the Matriculation Numbers of the three students in your team. Change the filename appropriately if you have less than three members in your team.**

```{r}
#install.packages("dplyr")
```


---


#### Summary of Data
```{r}
summary(retailData)
head(retailData)
tail(retailData)
```


---


#### Data Pre-processing 
```{r}
library(dplyr)
df<-filter(retailData,Description!="[A-Z].*")
df1<-filter(df,Description!="SAMPLES")
df2<-filter(df1,Quantity>0 )
df3<-filter(df2,UnitPrice>0)
df4<-filter(df3,Description!="Manual")
df5<-filter(df4,Description!="Adjust bad debt")
df6<-filter(df5,Description!="Next Day Carriage")
df7<-filter(df6,CustomerID!="NA")
df7<-filter(df7,Country != "Unspecified")
# write.csv(df7,'retaildataNew2.csv')
```

Some data pre-processing has been performed to handle inconsistent data, missing values and irrelevant data. 

* The `Description` with different format was found to be inconsistent with the stock code. (e.g. Stock code 20711 - JUMBO BAG TOYS, inconsistent name - lost in space). Since there were very few of these inconsistent data, and removing them would not affect the dataset significantly, the items with inconsistent description were removed.
* Samples are not very relevant in this analysis as they are freely given out to the customers when they purchase another product, as a marketing strategy. Also, the customer's decision to buy a product depends more heavily on their need for the product, rather than what samples come with it. Therefore, the samples would not affect the customer's purchase decision significantly, thus they were excluded from the dataset. 
* Those items with negative `Qantity` or `Unit Price` were also removed as it is impossible to have a product sold for negative quantity or unit price
* Manuals are add-ons to the products the customers purchase. Therefore, it is highly correlated to the product it is tied to, therefore does not add much information to the analysis. Thus, manuals were also removed from the dataset. 
* Items with `Description` such as "Adjust bad debt" or "Next Day Carriage" were also excluded as they are not product's names and are inconsistent with the dataset.
* Items with missing `CustomerID` were also removed from the dataset, as `CustomerID` is a key feature used for clustering. 
* Items with unspecified country were also removed as country of residence is important in understanding the purchase pattern of the customers. 


---


#### Removing Anomalies 

Removing anomalies is important as anomalies can affect the clustering. Removing anomalies can make the clustering process more efficient. 

```{r}
barplot(df7$Quantity, col = "blue")
```

From the barplot above, there are 2 anomalies of `Quantity` > 20000. These were removed.  

```{r}
df7<-filter(df7,Quantity<20000 )
barplot(df7$Quantity, col = "blue")
```


```{r}
barplot(df7$UnitPrice, col = "blue")
```

From the barplot above, there is an anomaly of `UnitPrice` > 3000. This data point was also removed.

```{r}
df7<-filter(df7,UnitPrice<3000 )
barplot(df7$UnitPrice, col = "blue")
```


---


#### Extracting 1000 most frequent items 

```{r}
t2 <- subset(df7, select = c("StockCode"))
countt2 <- table(unlist(t2))
countt2df<-as.data.frame(countt2)
colnames(countt2df) <- c("StockCode", "Freq")

countt2dfSort <- countt2df[rev(order(countt2df$Freq)),]

mostFreqProducts <- countt2dfSort[1:1000, "StockCode"]

mostFreqProductsExtracted <- df7[is.element(df7$StockCode, mostFreqProducts),]
head(mostFreqProductsExtracted) 

# write.csv(mostFreqProductsExtracted,'extracted.csv')
```

Using the `StockCode`, 1000 most popular items sold were extracted. The 1000 most popular items were used as they were probably generating the most revenue for the store, and it is logical to focus the marketing efforts for these most popular items. 


---


#### Creating Customer-Product matrix

```{r}
xtabs(Quantity ~ CustomerID + StockCode, data = mostFreqProductsExtracted, addNA = TRUE, sparse = TRUE)

```


```{r}
#install.packages("mclust")
```


---


#### Clustering

For clustering, we used K-Means Clustering algorithm. 

A 4 relevant columns were extracted for the analysis:
* `StockCode` : Stock code is essential to understand how many of which products are being sold.  
* `Quantity` : Quantity can help us to understand the purchase pattern customers have on each product. Some products would more likely to be sold as a bulk.
* `CustomerID` : Customer ID is also essential in understanding the customer's purchase pattern
* `Country` : Country is also important as customers from different countries might have different purchase patterns, for reasons such as differences in culture, GDP, preferences, etc. 


```{r}
clusterColumns <- subset(mostFreqProductsExtracted, select = c("StockCode", "Quantity", "CustomerID", "Country"))

summary(clusterColumns)
str(clusterColumns)
```

As K-Means clustering only works for numeric values, non-numeric values (`StockCode`, `Country`) were given numeric indexes for clustering purpose. 

```{r}
clusterColumns$numStockCode <- as.numeric(clusterColumns$StockCode)
clusterColumns$numCountry <- as.numeric(clusterColumns$Country)

summary(clusterColumns)

clusterColumns2 <- subset(clusterColumns, select = c("numStockCode", "Quantity", "CustomerID", "numCountry"))
```

K-Means Clusterng was performed with 2 features each. 


**1. Stock Code & Country**  

Features `StockCode` and `Country` were selected to understand what kind of products are sold frequently in which countries. Different numbers of clusters were experimented and plotted to determine the optimum number of clusters. Let K be the final number of clusters. 


```{r}
clusterColumns3 <- subset(clusterColumns, select = c("numStockCode", "numCountry"))
kMin <- 1
kMax <- 10
withinSS <- double(kMax - kMin + 1)
betweenSS <- double(kMax - kMin + 1)
for (K in kMin:kMax) {
  kMeansFit <-  kmeans(clusterColumns3, centers = K)
  withinSS[K] <- sum(kMeansFit$withinss)
  betweenSS[K] <- kMeansFit$betweenss
}
plot(kMin:kMax, betweenSS, pch=19, type="b", col="red",
     xlab = "Value of K", ylab = "Sum of Squares (Within and Between)")
points(kMin:kMax, withinSS, pch=19, type="b", col="green")
plot(kMin:kMax, withinSS, pch=19, type="b", col="green",
     xlab = "Value of K", ylab = "Sum of Squares (Within and Between)")
```

To evaluate the clusters, Within Sum of Squres and Between Sum of Squares were used. From the plots above, we can see that the Within Sum of Squares (Green line) drops drastically until K = 5 until almost 0 and plateus. At K = 5, the Between Sum of Squares (Red line) increases drastically and beyond K = 5 the value plateus. Therefore, the optimal number of clusters is 5. 

Therefore, 5 clusters were created. 

```{r}
K <- 5
kMeansFit <- kmeans(clusterColumns3, centers = K, nstart = 20)

# Note that the labels are stored as follows
clusterColumns3$cluster <- kMeansFit$cluster

```

These 5 clusters were then separated for deeper analysis. 

```{r}

mostFreqProductsExtracted$numStockCode <- clusterColumns$numStockCode
mostFreqProductsExtracted$numCountry <- clusterColumns$numCountry
mostFreqProductsExtracted$cluster <- clusterColumns3$cluster

#sort by cluster & write to file 

clsuterProductCountry <- mostFreqProductsExtracted[order(mostFreqProductsExtracted$cluster),]

# write.csv(clsuterProductCountry,'clusterProductCustomer.csv')

# separate each cluster into diff dataframe

allProductCountryClusters <- read.csv("clusterProductCustomer.csv", header = TRUE)

clusterLabels <- allProductCountryClusters$cluster

productCountryCluster1 <- allProductCountryClusters[clusterLabels == 1,]
productCountryCluster2 <- allProductCountryClusters[clusterLabels == 2,]
productCountryCluster3 <- allProductCountryClusters[clusterLabels == 3,]
productCountryCluster4 <- allProductCountryClusters[clusterLabels == 4,]
productCountryCluster5 <- allProductCountryClusters[clusterLabels == 5,]
```


* Cluster 1

Top 10 most frequently sold products from cluster1 was extracted. 

```{r}
cluster1 <- subset(productCountryCluster1, select = c("StockCode"))
clutercount1 <- table(unlist(cluster1))
count1<-as.data.frame(clutercount1)
colnames(count1) <- c("StockCode", "Freq")

sortcluster1 <- count1[rev(order(count1$Freq)),]

top10cluster1 <- sortcluster1[1:10, "StockCode"]

top10cluster1extracted <- productCountryCluster1[is.element(productCountryCluster1$StockCode, top10cluster1),]
```

```{r}
descriptions1 <- unique(top10cluster1extracted$Description)
descriptions1
price1 <- unique(top10cluster1extracted$UnitPrice)
price1

country1 <- unique(top10cluster1extracted$Country)
country1

```

There are some repetitions due to the spelling differences in the names (e.g. *JUMBO BAG VINTAGE DOILY*, *JUMBO BAG VINTAGE DOILEY*, *LUNCH BAG VINTAGE DOILY*, *LUNCH BAG VINTAGE DOILEY*), also different format (e.g. *SPOTTY BUNTING*, *BUNTING , SPOTTY*).

The most popular products in cluster 1 are affordable lunch wares that are popular all around the world. 


* Cluster 2

Top 10 most frequently sold products from cluster1 was extracted. 

```{r}
cluster2 <- subset(productCountryCluster2, select = c("StockCode"))
clutercount2 <- table(unlist(cluster2))
count2<-as.data.frame(clutercount2)
colnames(count2) <- c("StockCode", "Freq")

sortcluster2 <- count2[rev(order(count2$Freq)),]

top10cluster2 <- sortcluster2[1:10, "StockCode"]

top10cluster2extracted <- productCountryCluster2[is.element(productCountryCluster2$StockCode, top10cluster2),]
```

```{r}
descriptions2 <- unique(top10cluster2extracted$Description)
descriptions2

country2 <- unique(top10cluster2extracted$Country)
country2

price2 <- unique(top10cluster2extracted$UnitPrice)
price2

```

The most popular products in cluster 2 are antiques or items associated with tea culture (e.g. *ANTIQUE SILVER TEA GLASS ETCHED*, *ANTIQUE SILVER T-LIGHT GLASS*, *WOODEN FRAME ANTIQUE WHITE*,*60 TEATIME FAIRY CAKE CASES*). The countries in the cluster are mainly European countries which explains why the top products in the cluster are associated with tea. The unit prices across the cluster ranges from 0.40 (teaparty items) to 523.00 (antique silver tea glass).


* Cluster 3

Top 10 most frequently sold products from cluster1 was extracted. 

```{r}
cluster3 <- subset(productCountryCluster3, select = c("StockCode"))
clutercount3 <- table(unlist(cluster3))
count3<-as.data.frame(clutercount3)
colnames(count3) <- c("StockCode", "Freq")

sortcluster3 <- count3[rev(order(count3$Freq)),]

top10cluster3 <- sortcluster3[1:10, "StockCode"]

top10cluster3extracted <- productCountryCluster3[is.element(productCountryCluster3$StockCode, top10cluster3),]
```

```{r}
descriptions3 <- unique(top10cluster3extracted$Description)
descriptions3

country3 <- unique(top10cluster1extracted$Country)
country3

price3 <- unique(top10cluster3extracted$UnitPrice)
price3
```

The most popular products in cluster 3 are associated with kitchen items (*LUNCH BAG SUKI  DESIGN*,*REGENCY CAKESTAND 3 TIER*) that are relatively cheaper.

* Cluster 4

Top 10 most frequently sold products from the entire cluster was extracted. 

```{r}
cluster4 <- subset(productCountryCluster4, select = c("StockCode"))
clutercount4 <- table(unlist(cluster4))
count4<-as.data.frame(clutercount4)
colnames(count4) <- c("StockCode", "Freq")

sortcluster4 <- count4[rev(order(count4$Freq)),]

top10cluster4 <- sortcluster4[1:10, "StockCode"]

top10cluster4extracted <- productCountryCluster4[is.element(productCountryCluster4$StockCode, top10cluster4),]
```


```{r}

descriptions4 <- unique(top10cluster4extracted$Description)
descriptions4

price4 <- unique(top10cluster4extracted$UnitPrice)
price4

```

Cluster 4 containes some miscelleneous products(e.g. *VICTORIAN GLASS HANGING T-LIGHT*, *RETROSPOT TEA SET CERAMIC 11 PC*, *VINTAGE SNAP CARDS*). These products can also be used as small gifts. 


* Cluster 5

Top 10 most frequently sold products from the entire cluster was extracted. 

```{r}
cluster5 <- subset(productCountryCluster5, select = c("StockCode"))
clutercount5 <- table(unlist(cluster5))
count5<-as.data.frame(clutercount5)
colnames(count5) <- c("StockCode", "Freq")
sortcluster5 <- count5[rev(order(count5$Freq)),]

top10cluster5 <- sortcluster5[1:10, "StockCode"]

top10cluster5extracted <- productCountryCluster5[is.element(productCountryCluster5$StockCode, top10cluster5),]
```


```{r}
descriptions5 <- unique(top10cluster5extracted$Description)
descriptions5

price5 <- unique(top10cluster2extracted$UnitPrice)
price5

country5 <- unique(top10cluster5extracted$Country)
country5

```

Cluster 5 also contains some kitchen items, with a more varying price.   


---


#### Conclusion ####

UK is the country that appears in almost in all clusters. Since the biggest market for the shop is UK, the top selling products are related to their tea culture, from the cake cases to tea set. Other products that are sold world wide are often small affordable items, and we can see that the customer prefer to buy small affordable products online. Some high-end products are also popular, and this might be because buying those high-end items directly from the country of produce online is cheaper than buying the imported ones in the physical stores. 

The popularity of kitchen related items also signal that majority of the customers of this online shop are those who cook, and also those who pack lunch. 

Therfore, with this information, the marketing strategy could be adjusted to target customers in UK, and those who cook. The kitchen related items could also be promoted more, to attract more of the main group of customers. 